### 2016-03-16 paper/icfp16-submission/evaluation.tex
Made a pass over the complete paper, clarifying and breaking up some convoluted sentences, and removing references to "blind review" for ICFP.

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at our institution.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB \# hidden for blind review), % \#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml programs.
%
The second dataset --- widely used in the literature --- comes from a
similar course at the University of Washington~\cite{lerner_seminal:_2006},
from which we extracted 284 ill-typed programs.
%
Both datasets include programs that demonstrate a variety of functional
programming idioms, including (tail) recursive functions, higher-order
functions, polymorphic data types, and expression evaluators.

>>>>

\paragraph{Benchmarks}
We answer these questions on two sets of ill-typed programs, \ie
programs that were rejected by the \ocaml\ compiler because of a
type error.
%
The first dataset comes from the Spring 2014 undergraduate Programming
Languages course at our institution.
%
We recorded each interaction with the \ocaml\ top-level system over the
course of the first three assignments (IRB  
% \# hidden for blind review), 
\#140608),
from which we extracted \ucsdsize\ distinct, ill-typed \ocaml programs.
%
The second dataset --- widely used in the literature --- comes from a
similar course at the University of Washington~\cite{lerner_seminal:_2006},
from which we extracted 284 ill-typed programs.
%
Both datasets include programs that demonstrate a variety of functional
programming idioms, including (tail) recursive functions, higher-order
functions, polymorphic data types, and expression evaluators.

### 2016-03-16 paper/icfp16-submission/gallery2.tex
misc edits
Made a pass over the complete paper, clarifying and breaking up some convoluted sentences, and removing references to "blind review" for ICFP.

\section{The Advantage Of Traces}\label{sec:advantage-traces}

Next, we present a \emph{qualitative} evaluation that compares
the explanations provided by \toolname's dynamic witnesses with
the static reports produced by the \ocaml\ compiler and \sherrloc,
a state-of-the-art fault localization approach~\cite{zhang_toward_2014}.
%
In particular, we illustrate, using a series of examples drawn
from student programs in the \ucsdbench\ dataset how \toolname's
jump-compressed traces can get to the heart of the error by
%
highlighting the conflicting values that cause the program to get
stuck, rather that blaming a single one,
%
showing the steps necessary to reach the stuck state, and
%
not assuming that a function is correct just because it type-checks.
%
For each example we will present
(1)~the code,
(2)~the error message returned \ocaml,
(3)~the error locations returned by \ocaml\ (underlined) and \sherrloc\ (in bold)
\footnote{When the locations from \ocaml\ and \sherrloc\ overlap,
we just underline the relevant code.},
(4)~the jump-compressed trace produced by \toolname.

@@

  let _ = wwhile (__f__, 2)
\end{ecode}
%
Unfortunately, the student has forgotten to \emph{apply}
@f@ at all on line 2, and just matches it directly against
a pair.
This faulty definition of @wwhile@ still typechecks however,
and is assumed to be correct by and both \ocaml\
and \sherrloc\ which blame the use-site on line 10.
%
\begin{verbatim}
  This expression has type
    int -> int * bool
  but an expression was expected of type
    'a * bool
\end{verbatim}
%
\toolname\ synthesizes a trace that draws the eye to the
true error: the @match@ expression on line 2, and highlights
the conflict in matching a function against a pair pattern.
%
\begin{center}
  \includegraphics[height=150px]{wwhile.png}
\end{center}
%
By highlighting conflicting values (\ie\ the source and sink
of the problem) and not making any assumptions, \toolname\
focusses the user's attention on the piece of code that is
actually relevant to the error.

>>>>

\subsection{The Advantage Of Traces}\label{sec:advantage-traces}

Next, we present a \emph{qualitative} evaluation that compares
the explanations provided by \toolname's dynamic witnesses with
the static reports produced by the \ocaml\ compiler and \sherrloc,
a state-of-the-art fault localization approach~\cite{zhang_toward_2014}.
%
In particular, we illustrate, using a series of examples drawn
from student programs in the \ucsdbench\ dataset, how \toolname's
jump-compressed traces can get to the heart of the error. Our approach
%
highlights the conflicting values that cause the program to get
stuck, rather that blaming a single one,
%
shows the steps necessary to reach the stuck state, and
%
does not assume that a function is correct just because it type-checks.
%
For each example we will present
(1)~the code,
(2)~the error message returned \ocaml,
(3)~the error locations returned by \ocaml\ (underlined) and \sherrloc\ (in bold),\footnote{When the locations from \ocaml\ and \sherrloc\ overlap,
we just underline the relevant code.}
and (4)~the jump-compressed trace produced by \toolname.

@@

  let _ = wwhile (__f__, 2)
\end{ecode}
%
Unfortunately, the student has forgotten to \emph{apply}
@f@ at all on line 2, and just matches it directly against
a pair.
This faulty definition of @wwhile@ still typechecks however,
and is assumed to be correct by and both \ocaml\
and \sherrloc\ which blame the use-site on line 10.
%
\begin{verbatim}
  This expression has type
    int -> int * bool
  but an expression was expected of type
    'a * bool
\end{verbatim}
%
\toolname\ synthesizes a trace that draws the eye to the
true error: the @match@ expression on line 2, and highlights
the conflict in matching a function against a pair pattern.
%
\begin{center}
  \includegraphics[height=150px]{wwhile.png}
\end{center}
%
By highlighting conflicting values (\ie\ the source and sink
of the problem) and not making assumption about function correctness, \toolname\
focusses the user's attention on the piece of code that is
actually relevant to the error.

### 2016-03-16 paper/icfp16-submission/introduction.tex
Made a pass over the complete paper, clarifying and breaking up some convoluted sentences, and removing references to "blind review" for ICFP.

\paragraph{2. Visualizing Witnesses}
Our second contribution is an interactive visualization of the
execution of purely functional \ocaml\ programs, well-typed or not~(\S~\ref{sec:interactive}).
%
We extend the semantics to also build a \emph{reduction graph}
which records all of the small-step reductions and the context
in which they occur.
%
The graph lets us build a visualization that shows a sequence of
steps from the source witness to the stuck term and allows the user
to interactively expand the computation to expose intermediate steps,
by selecting an expression and choosing a traversal strategy.
%
The strategies include many of the standard debugging moves, \eg\
stepping \emph{forward} or \emph{into} or \emph{over} calls, as well
stepping or jumping \emph{backward} to understand how a particular
value was created, while preserving a context of the intermediate
steps that allow the user to keep track of a term's provenance.

>>>>

\paragraph{2. Visualizing Witnesses}
Our second contribution is an interactive visualization of the
execution of purely functional \ocaml\ programs, well-typed or not~(\S~\ref{sec:interactive}).
%
We extend the semantics to also build a \emph{reduction graph}
which records all of the small-step reductions and the context
in which they occur.
%
The graph lets us visualize the sequence of
steps from the source witness to the stuck term. The user can 
interactively expand the computation to expose intermediate steps
by selecting an expression and choosing a traversal strategy.
%
The strategies include many of the standard debugging moves, \eg\
stepping \emph{forward} or \emph{into} or \emph{over} calls, as well
stepping or jumping \emph{backward} to understand how a particular
value was created, while preserving a context of the intermediate
steps that allow the user to keep track of a term's provenance.
